---
title: "Covid-19 Data Analysis"
author: "Zach Barrett"
output: html_document
---
```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE, warning=FALSE, message=FALSE}
#load our data, make some initial adjustments
StateSummary <- data.frame(read.csv("JHCovidData.csv", header = TRUE ))
DetailedData <- data.frame(read.csv("JHCovidUS.csv", header = TRUE))
Census <- data.frame(read.csv("Census.csv", header = TRUE))
colnames(Census)[5] <- "Province_State"
```

## Case Fatality Rates

```{r, warning=FALSE, message=FALSE}
library(usmap)
library(ggplot2)
df <- StateSummary
colnames(df)[1] <- "state"

plot_usmap(data=df, values = "Case_Fatality_Ratio", color = "black")+
  scale_fill_continuous(
    low = "yellow", high = "red", name = "Case Fatality Ratio", label = scales::comma)+ 
  theme(legend.position = "right")+
  labs(title = "COVID-19 Case Fatality Rates")
```

# Population Density and Case Fatality Rates

```{r, warning=FALSE, message=FALSE}
library(usmap)
library(ggplot2)
require(gridExtra)
df <- StateSummary
colnames(df)[1] <- "state"
pop <- merge(x = StateSummary, y = Census, by = "Province_State")
colnames(pop)[1] <- "state"
pop$popDensity <-(pop$POPESTIMATE2019/pop$LandArea)

#Plot for Case Fatality Ratio
plot1 <- plot_usmap(data=df, values = "Case_Fatality_Ratio", color = "black")+
  scale_fill_continuous(
    low = "yellow", high = "red", name = "Case Fatality Ratio", label = scales::comma, limits=c(1.55,3))+ 
  theme(legend.position = "right")+
  labs(title = "COVID-19 Case Fatality Rates")

#Plot for Population Density
plot2 <- plot_usmap(data=pop, values = "popDensity", color = "black")+
  scale_fill_continuous(
    low = "pink", high = "purple", name = "People per square mi.", label = scales::comma, limits=c(90,1200))+
  theme(legend.position = "right")+
  labs(title = "Population Density")

grid.arrange(plot1, plot2, ncol=2)
```

# Is the above observation statistically significant?

```{r, warning=FALSE, message=FALSE}
df <- pop[is.finite(pop$popDensity),]
cor.test(df$popDensity, df$Case_Fatality_Ratio, use = "complete.obs")
```
# Hospital Beds and Case Fatality Rates

```{r, warning=FALSE, message=FALSE}
library(usmap)
library(ggplot2)
require(gridExtra)
df <- StateSummary
colnames(df)[1] <- "state"
pop <- merge(x = StateSummary, y = Census, by = "Province_State")
colnames(pop)[1] <- "state"
pop$bedsDensity <-(pop$StaffedBeds/pop$POPESTIMATE2019)

#Plot for Case Fatality Ratio
plot1 <- plot_usmap(data=df, values = "Case_Fatality_Ratio", color = "black")+
  scale_fill_continuous(
    low = "yellow", high = "red", name = "Case Fatality Ratio", label = scales::comma, limits=c(1.65,3))+
  theme(legend.position = "right")+
  labs(title = "COVID-19 Fatality Rates")

#Plot for Hospital Beds relative to population
plot2 <- plot_usmap(data=pop, values = "bedsDensity", color = "black")+
  scale_fill_continuous(
    low = "pink", high = "purple", name = "Beds per person", label = scales::comma, limits=c(.0025,.0035))+
  theme(legend.position = "right")+
  labs(title = "Total Hospital Beds per person")

grid.arrange(plot1, plot2, ncol=2)
```

# Is the above observation statistically significant?

```{r, warning=FALSE, message=FALSE}
cor.test(pop$bedsDensity, df$Case_Fatality_Ratio, use = "complete.obs")
```

# Covid-19 care by region

```{r, warning=FALSE, message=FALSE}
library(plotly)
library(ggplot2)
Census$REGION[which(Census$REGION == 1)] = "Northeast Region"
Census$REGION[which(Census$REGION == 2)] = "Midwest Region"
Census$REGION[which(Census$REGION == 3)] = "South Region"
Census$REGION[which(Census$REGION == 4)] = "West Region"
df <- merge(x = StateSummary, y = Census, by = "Province_State")
stored <- select(df, c("Province_State", "Deaths", "Confirmed", "POPESTIMATE2019", "REGION"))
stored <- filter(stored, REGION != "X")
stored$Death_By_Pop <- ((stored$Deaths)/(stored$POPESTIMATE2019)*1000)
stored$Cases_By_Pop <- ((stored$Confirmed)/(stored$POPESTIMATE2019)*1000)

p <-ggplot(stored, aes(x=Cases_By_Pop, y=Death_By_Pop, color = REGION))+
  geom_point()+
  geom_smooth(method=lm)+
  labs(title = "US COVID-19 Death and Case Rates by Region",
       x = "Cases per 1,000 people",
       y = "Deaths per 1,000 people")

ggplotly(p)
```

# Testing Rate and Death Rate

```{r, warning=FALSE, message=FALSE}
library(plotly)

names <- c("Rhode Island", "Massachusetts", "Alaska","New York","Vermont","District of Columbia","Illinois","North Dakota","Delaware","California","Maryland","New Mexico","West Virginia","Maine","Minnesota","Louisiana","Indiana","Colorado","Nebraska","New Jersey","Wisconsin","New Hampshire","Arizona","Wyoming","Montana","Michigan","Hawaii","South Carolina","Tennessee","Nevada","Ohio","Pennsylvania","Washington","Utah","Arkansas","Florida","Oklahoma","Missouri","Texas","Virginia","South Dakota","Georgia","Mississippi","Alabama","Kansas","Iowa","Idaho")

df <- merge(x = StateSummary, y = Census, by = "Province_State")
store <- select(df, c("Province_State", "Deaths", "Confirmed","Total_Test_Results", "POPESTIMATE2019"))
store <- filter(store, Province_State %in% names)
store$Death_By_Pop <- round(((store$Deaths/store$POPESTIMATE2019)), digits = 7)*1000
store$Tests_By_Pop <- round(((store$Total_Test_Results/store$POPESTIMATE2019)), digits = 5)*1000
store$Confirmed_By_Pop <- round(((store$Confirmed/store$POPESTIMATE2019)), digits = 5)*1000
store$Province_State <- factor(store$Province_State, levels = names)
store <- store[!(store$Province_State %in% c("Puerto Rico")), ]

plot_ly(store, x= ~Province_State, y= ~Tests_By_Pop, type = 'bar', name= 'Covid-19 Testing Rate') %>% 
  add_trace(x= ~Province_State, y= ~Death_By_Pop, mode = "markers", type = "scatter", yaxis = "y2", name = 'Covid-19 Death Rate') %>% 
  layout(
    title = "COVID-19 Testing and Death Rates by State", 
    xaxis = list(title = "State"), 
    yaxis = list(title = "Covid Testing Rate (per 1,000)", range = c(0, max(store$Tests_By_Pop, na.rm = TRUE)*1.05)),
    yaxis2 = list(
      title = "Death Rate (per 1,000)", 
      side = "right",
      overlaying = "y",
      showgrid = FALSE,
      range = c(0, max(store$Death_By_Pop, na.rm = TRUE)*1.05)
    )
  )
```

# Is the above observation statistically significant?

```{r, warning=FALSE, message=FALSE}
cor.test(store$Tests_By_Pop, store$Death_By_Pop)
```

# Testing rate and case rate

```{r, warning=FALSE, message=FALSE}
plot_ly(store, x= ~Province_State, y= ~Tests_By_Pop, type = 'bar', name= 'Covid-19 Testing Rate') %>% 
  add_trace(x= ~Province_State, y= ~Confirmed_By_Pop, mode = "markers", type = "scatter", yaxis = "y2", name = 'Covid-19 Case Rate') %>% 
  layout(
    title = "COVID-19 Testing and Case Rates by State", 
    xaxis = list(title = "State"), 
    yaxis = list(title = "Covid Testing Rate (per 1,000)", range = c(0, max(store$Tests_By_Pop, na.rm = TRUE)*1.05)),
    yaxis2 = list(
      title = "Case Rate (per 1,000)", 
      side = "right",
      overlaying = "y",
      showgrid = FALSE,
      range = c(0, max(store$Confirmed_By_Pop, na.rm = TRUE)*1.05)
    )
  )

```

# Is the above observation statistically significant?

```{r, warning=FALSE, message=FALSE}
cor.test(store$Tests_By_Pop, store$Confirmed_By_Pop)

# Does removing Hawaii improve our p-value?
store1 <- store[!(store$Province_State %in% c("Hawaii")), ]
cor.test(store1$Tests_By_Pop, store1$Confirmed_By_Pop, alternative = 'less')
```





